### 2021-11-22 input/proposal.tex
mostly on consistent partitioning...

At any time, a \process can decide to become a \emph{source}, hence
creating a new partition in the system by executing an \texttt{Add}
operation. This partition includes at least its source plus
neighboring \processes that estimate they are closer to this source
than any other one. Such a distance (or \emph{weight}) is
application-dependant: in the context of maintaining distributed
indexes, this would be about link latency that \nodes could monitor
by aggregating \texttt{ping}s.

\begin{definition}[\label{def:partitioning}Partitioning]
  Let $S \subseteq V$ be the set of \underline{s}ources, and $P_s$ be
  the \underline{p}artition including at least \Process~$s$, each
  \process belongs to at most one partition $\forall p,q \in V, \forall
  s_1,s_2 \in S: p \in P_{s_1} \wedge q \in P_{s_2} \implies p \neq q
  \vee s_1 = s_2$, and there exists at least one path $\pi_{ps}$ of
  \processes that belong to this partition $\forall q \in \pi_{ps}: q
  \in P_s$.
\end{definition}

@@

\begin{definition}[\underline{C}onsistent \underline{p}artitioning (CP)]
  Let $W_{xy} = W_{yx}$ be the positive \TODO{symmetric} \underline{w}eight
  between $x$ and $y$, $\Pi_{xz}$ be the shortest \underline{p}ath
  from $x$ to $z$ the weight of which $|\Pi_{xz}|$ is lower than any
  other path weight, with $|\Pi_{xx}|$ being the greatest lower bound
  of $x$, the only consistent partitioning $\mathcal{P}$ is a set of
  partitions $P_{s\in S}$ such that each \process belongs to a logical
  partition comprising its closest source: $\forall p \in P_{s_1}:
  \nexists P_{s_2}$ such that $|\Pi_{s_2p}| < |\Pi_{s_1p}|$.
\end{definition}

Unfortunately, \processes do not share a common global knowledge of
the network state. For \processes to reach consistent partitioning
together, a \Process $s$ \underline{a}dding a partition must send a
notification $\alpha_s$ to every \process that is closer to it than
any other source. Since epidemic dissemination and scoped broadcast
have well-aligned preoccupations, we assume implementations relying on
message forwarding from neighbor-to-neighbor.

\begin{theorem}[\label{theo:efb}\underline{E}ventual
  \underline{F}orwarding of \underline{B}est $(EFB) \implies CP$]
  Assuming reliable communication links where a correct \process $q$
  eventually receives the message $m$ \underline{s}ent to it by a
  \process $p$ ($s_{pq}(m) \implies r_{q}(m)$), \processes eventually
  reach consistent partitioning if each \process eventually
  \underline{f}orwards
  ($f_p(m) \implies \forall \langle p, q\rangle \in E: s_{pq}(m)$) its
  best known partition. \TODO{setminus parent to comply with implem'.}
\end{theorem}

@@

\begin{proof}
  When a \process $s_1$ becomes a source, it belongs to its own
  partition, for there exists no better partition than its own:
  $\forall p \in V: |\Pi_{s_1 s_1}| < |\Pi_{s_1 p}|$. It delivers,
  hence forwards such an $\alpha$ message to its neighbors. Since
  communication links are reliable, neighboring \processes eventually
  receive such a notification $\forall \langle s_1, q \rangle \in E,
  s_1 \in S \iff \eventually r_q(\alpha_{s_1}^{w_{s_1 q}})$. Most
  importantly, whatever the order of received messages, every \process
  $q'$ in this neighborhood -- such that there exists no better
  partition $s_2$ than the received one -- delivers and forwards it:
  $\forall s_2 \in S: |\Pi_{q' s_1}| < |\Pi_{q' s_2}| \implies
  \eventually d_{q'}(\alpha_{s_1}^{w_{s_1 q'}})$. By transitivity, the
  message originating from $s_1$ reaches all such \processes through
  their shortest paths: $\forall q'' \in V, s_1, s_2 \in S: |\Pi_{q''
    s_1}| < |\Pi_{q'' s_2}| \implies \eventually
  d_{q''}(\alpha_{s_1}^{|\Pi_{s_1 q''}|})$.  Since there exists only
  one best sum of weights per \process that can never be retracted,
  \processes eventually reach consistent partitioning.
  %% /!\ not equivalence for there exists other implementations.
  %% \item [$CP \implies BEF$:] By contradiction, if a \process $q \in
  %%   \Pi_{sp} = [s, \ldots, q, q', \ldots, p]$ with $s, q, p \in P_s$
  %%   does not forward its received $\alpha_s^{|\Pi_{sq}|}$, then
  %%   following \processes from $q'$ to $p$ may mistake another
  %%   partition for their  because it needs the weight $W_{pq}$.
  %% \end{asparadesc}
\end{proof}

@@

Algorithm~\ref{algo:add} shows the instructions that implement
consistent partitioning when
\begin{inparaenum}[(i)]
\item weights are scalar values,
\item \processes only add new partitions to the system,
\item and \processes never crash nor leave the system.
\end{inparaenum}
Figure~\ref{fig:add} illustrates its behavior on a system comprising 4
\processes $a$, $b$, $c$, and $d$. \Process~$a$ and \Process~$d$
become the sources of their partition. They \NAMEB a notification
\underline{a}dd message: $\alpha_a^0$ and $\alpha_d^0$. They
initialize their own state with the lowest possible bound $0$ (see
Line~\ref{line:lowestbound}), and send a message to each of their
neighbors by accumulating the corresponding edge weight (see
Line~\ref{line:accumulator}). In Figure~\ref{fig:addC}, \Process~$b$
receives $\alpha_{d}^{1}$. Since it improves its own partition
distance, it keeps it and forwards it to its neighbors. In
Figure~\ref{fig:addD}, \Process~$b$ discards $\alpha_{a}^{2}$, for it
does not improve its partition distance. \Processes $c$ and $d$ will
never acknowledge that Source~$a$ exists. Ultimately, \processes
discard last transiting messages. Despite the obvious lack of traffic
optimization, the system reaches consistent partitioning.

@@

A naive implementation resembles echos in acoustics:
\underline{d}elete notifications $\delta$ propagate as long as
receivers' current partition is the targeted one, and when delete
notifications reach the bordering \processes of the deleted partition,
it triggers a competition that goes backward to fill the gap left open
by removals.  For example, in Figure~\ref{fig:del}, two partitions
initially exist: $P_a$ and $P_d$ that respectively include $\{a\}$,
and $\{b, c, d\}$. In Figure~\ref{fig:delA}, \Process~$a$ deletes its
partition. It notifies all neighboring \processes~--~here only
\Process~$b$~-- that may belong to its partition using \NAMEB. Upon
receipt, \Process~$b$ discards the delete notification $\delta_a$
since $\delta_a$ does not target the former's partition
$P_d$. \Process~$b$ sends its own best partition $\alpha_d^3$ that may
be the best for \Process~$a$. Eventually, every \process belongs to
Partition $P_d$. In this scenario, they reach consistent partitioning.

@@

\begin{definition}[\label{def:purge}Message purge]
  A system \underline{p}urges a message $\mathcal{P}(m)$ when every
  \node $q$ that delivered $m$ eventually delivers another message
  $m'$ never followed by the delivery of the former message $m$:
  $d_q(m) \implies \exists m': \eventually (d_q(m)
  \rightarrow d_q(m') \wedge \neg (d_q(m') \rightarrow d_q(m)))$.
\end{definition}

>>>>

At any time, a \process can decide to become a \emph{source}, hence
creating a new partition in the system by executing an \texttt{Add}
operation. This partition includes at least its source plus
neighboring \processes that estimate they are closer to this source
than any other one. Such a distance (or \emph{weight}) is
application-dependant: in the context of maintaining distributed
indexes, this would be about link latency that \nodes could monitor by
aggregating \texttt{ping}s; or operational costs when dealing with
multiple tenants.

\begin{definition}[\label{def:partitioning}Partitioning]
  Let $S \subseteq V$ be the set of \underline{s}ources, and $P_{s\in
    S}$ be the \underline{p}artition including at least \Process~$s$,
  each \process belongs to at most one partition $\forall p,q \in V,
  \forall s,s' \in S: (p \in P_{s} \wedge q \in P_{s'}) \implies (p \neq
  q \vee s = s')$, and there exists at least one path $\pi_{ps}$ of
  \processes that belong to this partition $\forall q \in \pi_{ps}: q
  \in P_s$.
\end{definition}

@@

\begin{definition}[\underline{C}onsistent \underline{p}artitioning (CP)]
  Let $W_{xy}$ be the positive \underline{w}eight from $x$ to $y$,
  $\Pi_{xz}$ be the shortest \underline{p}ath from $x$ to $z$ the
  weight of which $|\Pi_{xz}|$ is lower than any other path weight,
  with $|\Pi_{xx}|$ being the greatest lower bound of $x$, the only
  consistent partitioning $\mathcal{P}$ is a set of partitions
  $P_{s\in S}$ such that each \process belongs to a logical partition
  comprising its closest source: $\forall p \in P_{s}: \nexists
  P_{s'}$ such that $|\Pi_{s'p}| < |\Pi_{sp}|$.
\end{definition}

Unfortunately, \processes do not share a common global knowledge of
the network state. For \processes to reach consistent partitioning
together, a \Process $s$ \underline{a}dding a partition must send a
notification $\alpha_s$ to every \process that is closer to it than
any other source. Since epidemic dissemination and scoped broadcast
have well-aligned objectives, we assume implementations relying on
message forwarding from neighbor-to-neighbor.

\begin{definition}[\label{def:forwarding}Forwarding]
  A \process $p$ \underline{f}orwards a message $m$ by
  \underline{s}ending it to its neighbors: $f_p(m) \implies \forall
  \langle p, q\rangle \in E: \TODO{s_{pq}(m)}$.
\end{definition}

\begin{theorem}[\label{theo:efb}\underline{F}orwarding of \underline{B}est (FB)
  $\implies\eventually$CP] Assuming reliable communication links
  ($s_{pq}(m) \iff \eventually r_{qp}(m)$), \processes eventually
  reach consistent partitioning if each \process delivers the message
  notifying its shortest path weight to a source among its set of
  \underline{r}eceived messages $R_p$ ($(m = \min R_p) \implies
  d_p(m)$), and forwards the message notifying its shortest path
  weight to a source $s$ among its set of \underline{d}elivered
  messages $D_p$ ($(m = \min D_p) \implies f_p(m)$).
\end{theorem}

@@

\begin{proof}
  When there are no sources, no \process receives, nor delivers hence
  forwards any message. \Processes do not belong to any partition.

  \noindent Whenever a \process $s$ becomes a source, it delivers its
  own message $d_p(\alpha_{s})$. Whatever its set of received
  messages, its own partition $P_s$ becomes and remains its closest
  partition $\alpha_{s} = \min D_p$ since $\forall p \neq s: |\Pi_{s
    s}| < |\Pi_{s p}|$.

  \noindent Such a source $s$ forwards $\alpha_{s}$ to its neighbors.
  Since communication links are reliable, neighboring \processes
  eventually receive such a notification $\forall \langle s, q \rangle
  \in E \iff \eventually r_q(\alpha_{s})$. Most importantly,
  whatever the order of received messages, every \process $q$ in this
  neighborhood such that there exists no closer source $s'$ than the
  received one $s$, delivers it, since $|\Pi_{sq}| < |\Pi_{s'q}|$.

  \noindent They forward it and their respective neighbors eventually
  receive it. By transitivity, the message originating from $s$
  reaches all \processes belonging to $P_s$ at least through their
  shortest paths: $\forall q' \in V, s, s' \in S: |\Pi_{s q'}| <
  |\Pi_{s' q'}| \implies \eventually d_{q'}(\alpha_{s})$.  Since there
  exists only one best sum of weights per \process that can never be
  retracted, \processes eventually reach consistent partitioning.
  %% /!\ not equivalence for there exists other implementations.
  %% \item [$CP \implies BEF$:] By contradiction, if a \process $q \in
  %%   \Pi_{sp} = [s, \ldots, q, q', \ldots, p]$ with $s, q, p \in P_s$
  %%   does not forward its received $\alpha_s^{|\Pi_{sq}|}$, then
  %%   following \processes from $q'$ to $p$ may mistake another
  %%   partition for their  because it needs the weight $W_{pq}$.
  %% \end{asparadesc}
\end{proof}

@@

Algorithm~\ref{algo:add} shows the instructions that implement
forwarding of best to reach consistent partitioning when
\begin{inparaenum}[(i)]
\item weights are scalar values,
\item \processes only add new partitions to the system,
\item and \processes never crash nor leave the system.
\end{inparaenum} \TODO{Disambiguator and comparison with $\alpha < A$.}
\TODO{Why not towards parent?}
Figure~\ref{fig:add} illustrates its behavior on a system comprising 4
\processes $a$, $b$, $c$, and $d$. \Process~$a$ and \Process~$d$
become the sources of their partition. They \NAMEB a notification
\underline{a}dd message: $\alpha_a^0$ and $\alpha_d^0$. They
initialize their own state with the lowest possible bound $0$ (see
Line~\ref{line:lowestbound}), and send a message to each of their
neighbors by accumulating the corresponding edge weight (see
Line~\ref{line:accumulator}). In Figure~\ref{fig:addC}, \Process~$b$
receives $\alpha_{d}^{1}$. Since it improves its own partition
distance, it keeps it and forwards it to its neighbors. In
Figure~\ref{fig:addD}, \Process~$b$ discards $\alpha_{a}^{2}$, for it
does not improve its partition distance. \Processes $c$ and $d$ will
never acknowledge that Source~$a$ exists. Ultimately, \processes
discard last transiting messages. Despite the obvious lack of traffic
optimization, the system reaches consistent partitioning.

@@

A naive implementation resembles echos in acoustics:
\underline{d}elete notifications $\delta$ propagate as long as
receivers' current partition is the targeted one, and when delete
notifications reach the bordering \processes of the deleted partition,
it triggers a competition -- or an echo -- that goes backward to fill
the gap left open by removals.  For example, in Figure~\ref{fig:del},
two partitions initially exist: $P_a$ and $P_d$ that respectively
include $\{a\}$, and $\{b, c, d\}$. In Figure~\ref{fig:delA},
\Process~$a$ deletes its partition. It notifies all neighboring
\processes~--~here only \Process~$b$~-- that may belong to its
partition using \NAMEB. Upon receipt, \Process~$b$ discards the delete
notification $\delta_a$ since $\delta_a$ does not target the former's
partition $P_d$. \Process~$b$ sends its own best partition
$\alpha_d^3$ that may be the best for \Process~$a$. Eventually, every
\process belongs to Partition $P_d$. In this scenario, they reach
consistent partitioning.

@@

\begin{definition}[\label{def:purge}Message purge]
  A system \underline{p}urges a message $\TODO{\mathcal{P}}(m)$ when every
  \node $q$ that delivered $m$ eventually delivers another message
  $m'$ never followed by the delivery of the former message $m$:
  $d_q(m) \implies \exists m': \eventually (d_q(m)
  \rightarrow d_q(m') \wedge \neg (d_q(m') \rightarrow d_q(m)))$.
\end{definition}
